---
output:
  pdf_document: default
  html_document: default
---
```{r}
library(vegan)
data(BCI)
## UTM Coordinates (in metres)
BCI_xy = data.frame(x = rep(seq(625754, 626654, by=100), each=5), 
                    y = rep(seq(1011569,  1011969, by=100), len=50))
```

^^Spacial coordianates of the quadrat
```{R}
BCI
```

```{r}
BCI_common<- BCI$Faramea.occidentalis
BCI_rare<- BCI$Eugenia.nesiotica
```

```{r}
hist(BCI_common)
hist(BCI_rare)
```

```{r}
col_brks = hist(BCI_common, plot=F)$breaks
col_indices = as.numeric(cut(BCI_common, col_brks))
cols = rev(terrain.colors(length(col_brks)))
plot(BCI_xy, cex=2, pch=19, col=cols[col_indices])
```
```{r}
col_brks = hist(BCI_rare, plot=F)$breaks
col_indices = as.numeric(cut(BCI_rare, col_brks))
cols = rev(terrain.colors(length(col_brks)))
plot(BCI_xy, cex=2, pch=19, col=cols[col_indices])
```





```{r}
BCI_SR= apply(BCI, 1, function(x) sum(x > 0))
hist(BCI_SR)
```


```{r}
common_dist = dist(BCI_common)
xy_dist = dist(BCI_xy)
max_dist = max(xy_dist) / 2
plot(xy_dist, common_dist)
abline(lm(common_dist ~ xy_dist), lwd=3, col='red')
lines(lowess(xy_dist, sr_dist), lwd=3, col='pink')
abline(v = max_dist, col='red', lwd=3, lty=2)
```

```{r}
rare_dist = dist(BCI_rare)
xy_dist = dist(BCI_xy)
max_dist = max(xy_dist) / 2
plot(xy_dist, rare_dist)
abline(lm(rare_dist ~ xy_dist), lwd=3, col='red')
lines(lowess(xy_dist, sr_dist), lwd=3, col='pink')
abline(v = max_dist, col='red', lwd=3, lty=2)
```

```{R}
obs_cor = cor(xy_dist, common_dist)
obs_cor
```

```{R}
obs_cor = cor(xy_dist, rare_dist)
obs_cor
```

```{r}
nperm = 1000
null_cor = obs_cor
for (i in 2:nperm) {
    # shuffle the rows of the spatial coordinates
    tmp_xy = BCI_xy[sample(nrow(BCI_xy)), ]
    # correlation between the shuffled spatial coordinates and sr_dist
    null_cor[i] = cor(dist(tmp_xy), common_dist)
}
# compute the p-value
sum(null_cor >= obs_cor) / nperm 
```

```{r}
nperm = 1000
null_cor = obs_cor
for (i in 2:nperm) {
    # shuffle the rows of the spatial coordinates
    tmp_xy = BCI_xy[sample(nrow(BCI_xy)), ]
    # correlation between the shuffled spatial coordinates and sr_dist
    null_cor[i] = cor(dist(tmp_xy), rare_dist)
}
# compute the p-value
sum(null_cor >= obs_cor) / nperm 
```

```{r}
common_mantel = mantel(xy_dist, common_dist)
common_mantel
```

```{r}
rare_mantel = mantel(xy_dist, rare_dist)
rare_mantel
```
```{r}
boxplot(list(null_cor, common_mantel$perm), horizontal = T, boxwex = 0.5,
        names = c('mine', 'theirs'), xlab='Correlation')
abline(v=obs_cor, col='red')
```

```{r}
boxplot(list(null_cor, rare_mantel$perm), horizontal = T, boxwex = 0.5,
        names = c('mine', 'theirs'), xlab='Correlation')
abline(v=obs_cor, col='red')
```

```{r}
comm_dist = vegdist(BCI)
plot(xy_dist, comm_dist)
abline(lm(comm_dist ~ xy_dist), lwd=3, col='red')
lines(lowess(xy_dist, comm_dist), lwd=3, col='pink')
lines(lowess(xy_dist, comm_dist, f=0.1), lwd=3, col='blue')

abline(v = max_dist, col='red', lwd=3, lty=2)
```


Both the rare and the common species are not spacially dependent. The permutation test for both the rare and common species had p values greater than 0.05. The mantel test also showed no significant relationship between species (both common and rare) and spacial dependancy. 










```{r}
sp_ids = c("Cordia.lasiocalyx", "Hirtella.triandra",
           "Picramnia.latifolia", "Quassia.amara",
           "Tabernaemontana.arborea", "Trattinnickia.aspera", 
           "Xylopia.macrantha")
```

```{r}
sp_a<- BCI$Cordia.lasiocalyx
sp_b<- BCI$Hirtella.triandra
sp_c<- BCI$Picramnia.latifolia
sp_d<- BCI$Quassia.amara
sp_e<- BCI$Tabernaemontana.arborea
sp_f<- BCI$Trattinnickia.aspera
sp_g<- BCI$Xylopia.macrantha
sp_one<- BCI$Drypetes.standleyi
```

```{r}
sp_a
```

```{r}
sp_c
```

```{r}
comm_mantel = mantel(xy_dist, comm_dist)
comm_mantel
```

```{r}
common_dist22 = dist(sp_a)
rare_dist22 = dist(sp_c)
```

```{r}
sr_corlog = mantel.correlog(common_dist22, xy_dist)
comm_corlog = mantel.correlog(common_dist22, xy_dist)
sr_corlog
comm_corlog
```

```{r}
sr_corlog_rare = mantel.correlog(rare_dist22, xy_dist)
comm_corlog_rare = mantel.correlog(rare_dist22, xy_dist)
sr_corlog_rare
comm_corlog_rare
```

```{r}
par(mfrow=c(1,2))
plot(sr_corlog)
mtext(side=3, 'Species Richness')
abline(v = max_dist, col='red', lwd=3, lty=2)
plot(comm_corlog)
mtext(side=3, 'Community Composition')
abline(v = max_dist, col='red', lwd=3, lty=2)
```


```{r}
par(mfrow=c(1,2))
plot(sr_corlog_rare)
mtext(side=3, 'Species Richness')
abline(v = max_dist, col='red', lwd=3, lty=2)
plot(comm_corlog_rare)
mtext(side=3, 'Community Composition')
abline(v = max_dist, col='red', lwd=3, lty=2)
```

```{r}
x <- BCI_xy$x
y <- BCI_xy$y

two_dat = data.frame(BCI, BCI_xy)
two_lm = gls(sp_a ~ sp_c, data=BCI)
plot(Variogram(two_lm, form= ~ x + y))
```

```{r}
res = residuals(two_lm)
plot(dist(two_dat[, c('x', 'y')]), dist(res))
lines(lowess(dist(two_dat[, c('x', 'y')]), dist(res)), col='red', lwd=2)
abline(v = max_dist, col='red', lwd=3, lty=2)
```

```{r}
two_exp111 = update(two_lm, corr=corExp(form=~x + y))
plot(Variogram(two_exp111, maxDist = max_dist))
```



```{R}
two_rat_nug = update(two_lm, corr=corRatio(form=~x + y, nugget=T))
plot(Variogram(two_rat_nug, maxDist = max_dist))
```

```{r}
plot(Variogram(two_rat_nug, resType='n', maxDist = max_dist))
```

```{r}
all<-anova(two_lm, two_exp111, two_rat_nug, test=F)
all
```




Including the spatial error does not have a big impact on the model. Adding in the spatial error did not significantly impact the AIC value AKA improve the model. 

Adding in the spatial error may not have made a large impact because spatial dependancy was not signficant before either. Therefore, this could be why adding in spatial error may not have made a large impact on the model. 


